RIScatter: Unifying Backscatter Communication and Reconfigurable Intelligent Surface

### 2023-02-24 main.tex
:construction: Update contribution

\begin{abstract}
	% \gls{bc} nodes harvest energy from and modulate information over an external electromagnetic wave by manipulating the magnitude, phase, and/or frequency of the scattered signal.
	% \gls{sr} incorporates a passive scatter node into existing radio networks to create additional propagation path and ride its own information towards the cooperative receiver.
	% \gls{ris} is a programmable reflector array that adapts the phase shift response to enhance or suppress signal strength in specific directions.
	\gls{bc} nodes harvest energy from and modulate information over an external electromagnetic wave.
	\gls{ris} adapts its phase shift response to enhance or attenuate channel strength in specific directions.
	In this paper, we show how those two seemingly different technologies (and their derivatives) can be unified into a single architecture called RIScatter.
	RIScatter consists of multiple dispersed or co-located scatter nodes, whose reflection states are adapted to partially modulate their own information and partially engineer the wireless channel.
	% This contrasts with \gls{bc} (resp. \gls{ris}) where the reflection pattern is exclusively a function of the information symbol (resp. \gls{csi}).
	The key principle is to render the probability distribution of reflection states as a joint function of the information source, \gls{csi}, and \gls{qos} of coexisting active primary and passive backscatter links.
	This enables RIScatter to softly bridge \gls{bc} and \gls{ris}; boil down to either under specific input distribution; or evolve in a mixed form for heterogeneous traffic control and universal hardware design.
	% To accommodate the characteristics of active-passive coexisting networks, we propose a practical \gls{sic}-free receiver
	To reap the benefits of RIScatter, we also propose a co-located \gls{sic}-free receiver that semi-coherently decodes the backscatter information, recovers the reflection pattern, and coherently decodes the primary link.
	For a single-user multi-node RIScatter network, we characterize the achievable primary-(total-)backscatter rate region by designing the input distribution at scatter nodes, the active beamforming at the \gls{ap}, and the energy detection regions at the user.
	Simulation results demonstrate RIScatter nodes can exploit the scattered paths to smoothly shift between backscatter modulation and passive beamforming.
\end{abstract}

@@

	% \begin{subsection}{Contributions}
	% \emph{First,} we propose RIScatter as a novel scatter protocol that unifies \gls{bc} and \gls{ris} by adaptive reflection state (backscatter input) distribution design.
	% \emph{First,} we propose RIScatter as a novel scatter protocol that simultaneously modulates the nodes information and engineers the wireless channel via adaptive reflection state (backscatter input) distribution design.
	\emph{First,} we propose RIScatter as a novel scatter protocol that simultaneously modulates the node information and engineers the wireless channel via adaptive reflection state (backscatter input) distribution design.
	The concept is shown in Fig.~\subref*{fg:riscatter} where the message of one or more RIScatter nodes and a primary transmitter are sequentially decoded by a co-located receiver.
	% The reflection pattern is random with the guidance of input distribution as a joint function of the information source, \gls{csi}, and \gls{qos}
	% The reflection pattern is a discrete random variable with probability mass function
	% (co-located or dispersed)
	Each reflection state simultaneously acts as information codeword and passive beamforming codeword.
	The reflection pattern of each node is randomly chosen with the guidance of reflection state distribution, which is designed as a joint function of the information source, \gls{csi}, and \gls{qos}.
	% That is, each reflection state simultaneously acts as information codeword and passive beamforming codeword (symbol-level passive precoding).
	% Since each reflection state simultaneously acts as information codeword and passive beamforming codeword (symbol-level passive precoding).
	Such an adaptive channel coding boils down to the degenerate distribution of \gls{ris} when the primary link is prioritized, and outperforms the uniform distribution of line-coded \gls{bc} (by accounting the \gls{csi}) when the backscatter link is prioritized.
	% Joint and independent encoding are available when multiple RIScatter nodes are
	Multiple RIScatter nodes can be co-located with joint encoding or dispersed with independent encoding.
	% The reflection state is neither fully random (cf. \gls{bc}) nor fully deterministic (cf. \gls{ris}), but semi-random in reflection state distribution.
	% RIScatter exploits the transmitter-scatter-receiver path for simultaneous backscatter modulation and dynamic passive beamforming, and can smoothly transition in between under different \gls{qos}.
	% A comparison of RIScatter to \gls{bc} and \gls{ris} is summarized in Table~\ref{tb:scattering_applications}.
	Table~\ref{tb:scattering_applications} compares RIScatter to \gls{bc} and \gls{ris}.
	However, two major challenges for RIScatter are the cooperative receiver design and the backscatter input distribution design.
	% such a novel scatter protocol encounters two major challenges, namely the cooperative receiver design and the adaptive backscatter input distribution design.
	This is the first paper to unify \gls{bc} and \gls{ris} from the perspective of input distribution.

	\emph{Second,} we address the first challenge and propose a practical receiver that accommodates the properties of active-passive coexisting networks.
	It semi-coherently decodes the backscatter information by energy detection, recovers the exact reflection pattern, and coherently decodes the primary link.
	% Double modulation enables symbol-level passive precoding such that the
	Once successfully decoded, the multiplicative backscatter interference can be modelled as dynamic passive beamforming (rather than being re-precoded and cancelled), and thus the energy detection becomes part of channel training.
	The proposed receiver significantly improves backscatter throughput since it is suitable for arbitrary input distribution and spreading factor.
	This is the first paper to consider practical backscatter-primary decoding and treat semi-coherent backscatter detection as part of primary channel training.

	\emph{Third,} we address the second challenge by considering a specific scenario where multiple co-located/dispersed RIScatter nodes ride over an active point-to-point \gls{miso} transmission to perform backscatter modulation and dynamic passive beamforming towards a nearby user.
	For a given \gls{qos}, we first express the primary and total backscatter rates as functions of \emph{input distribution} at RIScatter nodes, \emph{active beamforming} at the \gls{ap}, and \emph{backscatter decision regions} at the user.
	Then, we characterize the achievable rate region by optimizing those variables under different \gls{qos}.
	Since the original problem is highly non-convex, we propose a suboptimal \gls{bcd} algorithm where the \gls{kkt} input distribution is numerically evaluated by limit of sequences, the active beamforming is iteratively updated by \gls{pgd} accelerated by \gls{bls}, and the decision regions are refined by existing sequential quantization methods for \gls{dmtc}.
	Uniquely, we account for \gls{csi}, \gls{qos}, and backscatter constellation in those optimizations, and the proposed backscatter input distribution design is applicable to more general scenarios.
	This is also the first paper to reveal the importance of backscatter input distribution and decision region designs in active-passive coexisting networks.

>>>>

\begin{abstract}
	% \gls{bc} nodes harvest energy from and modulate information over an external electromagnetic wave by manipulating the magnitude, phase, and/or frequency of the scattered signal.
	% \gls{sr} incorporates a passive scatter node into existing radio networks to create additional propagation path and ride its own information towards the cooperative receiver.
	% \gls{ris} is a programmable reflector array that adapts the phase shift response to enhance or suppress signal strength in specific directions.
	\gls{bc} nodes harvest energy from and modulate information over an external electromagnetic wave.
	\gls{ris} adapts its phase shift response to enhance or attenuate channel strength in specific directions.
	In this paper, we show how those two seemingly different technologies (and their derivatives) can be unified into a single architecture called RIScatter.
	RIScatter consists of multiple dispersed or co-located scatter nodes, whose reflection states are adapted to partially modulate their own information and partially engineer the wireless channel.
	% This contrasts with \gls{bc} (resp. \gls{ris}) where the reflection pattern is exclusively a function of the information symbol (resp. \gls{csi}).
	The key principle is to render the probability distribution of reflection states as a joint function of the information source, \gls{csi}, and \gls{qos} of coexisting active primary and passive backscatter links.
	This enables RIScatter to softly bridge \gls{bc} and \gls{ris}; boil down to either under specific input distribution; or evolve in a mixed form for heterogeneous traffic control and universal hardware design.
	% To accommodate the characteristics of active-passive coexisting networks, we propose a practical \gls{sic}-free receiver
	To reap the benefits of RIScatter, we also propose a co-located \gls{sic}-free receiver that semi-coherently decodes the backscatter information, recovers the reflection pattern (and composite channel), and coherently decodes the primary link.
	For a single-user multi-node RIScatter network, we characterize the achievable primary-(total-)backscatter rate region by designing the input distribution at scatter nodes, the active beamforming at the \gls{ap}, and the energy detection regions at the user.
	Simulation results demonstrate RIScatter nodes can exploit the scattered paths to smoothly shift between backscatter modulation and passive beamforming.
\end{abstract}

@@

	% \begin{subsection}{Contributions}
	% \emph{First,} we propose RIScatter as a novel scatter protocol that unifies \gls{bc} and \gls{ris} by adaptive reflection state (backscatter input) distribution design.
	% \emph{First,} we propose RIScatter as a novel scatter protocol that simultaneously modulates the nodes information and engineers the wireless channel via adaptive reflection state (backscatter input) distribution design.
	\emph{First,} we propose RIScatter as a novel scatter protocol that simultaneously modulates the node information and engineers the wireless channel via adaptive reflection state (backscatter input) distribution design.
	The concept is shown in Fig.~\subref*{fg:riscatter} where the message of one or more RIScatter nodes and a primary transmitter are sequentially decoded by a co-located receiver.
	% The reflection pattern is random with the guidance of input distribution as a joint function of the information source, \gls{csi}, and \gls{qos}
	% The reflection pattern is a discrete random variable with probability mass function
	% (co-located or dispersed)
	Each reflection state simultaneously acts as an information codeword and a passive beamforming codeword.
	The reflection pattern of each node is randomly chosen with the guidance of input distribution that is designed as a joint function of the information source, \gls{csi}, and \gls{qos}.
	% That is, each reflection state simultaneously acts as information codeword and passive beamforming codeword (symbol-level passive precoding).
	% Since each reflection state simultaneously acts as information codeword and passive beamforming codeword (symbol-level passive precoding).
	Such an adaptive channel coding boils down to the degenerate distribution of \gls{ris} when the primary link is prioritized, and outperforms the uniform distribution of line-coded \gls{bc} (by accounting the \gls{csi}) when the backscatter link is prioritized.
	% Joint and independent encoding are available when multiple RIScatter nodes are
	Multiple RIScatter nodes can be co-located with joint encoding or dispersed with independent encoding.
	% The reflection state is neither fully random (cf. \gls{bc}) nor fully deterministic (cf. \gls{ris}), but semi-random in reflection state distribution.
	% RIScatter exploits the transmitter-scatter-receiver path for simultaneous backscatter modulation and dynamic passive beamforming, and can smoothly transition in between under different \gls{qos}.
	% A comparison of RIScatter to \gls{bc} and \gls{ris} is summarized in Table~\ref{tb:scattering_applications}.
	Table~\ref{tb:scattering_applications} compares RIScatter to \gls{bc} and \gls{ris}.
	However, two major challenges for RIScatter are the cooperative receiver design and the backscatter input distribution design.
	% such a novel scatter protocol encounters two major challenges, namely the cooperative receiver design and the adaptive backscatter input distribution design.
	This is the first paper to unify \gls{bc} and \gls{ris} from the perspective of input distribution.

	\emph{Second,} we address the first challenge and propose a practical receiver that semi-coherently decodes the backscatter information, recovers the exact reflection pattern (and composite channel), and coherently decodes the primary link.
	% Energy detection is used to decode the multiplicative backscatter information under primary uncertainty, and disjoint decision regions are formulated
	% Energy detection is used to decode the multiplicative backscatter information under primary uncertainty, and it divides the accumulated energy domain into disjoint decision regions to construct \gls{dmc}
	% Energy detection is used for backscatter decoding that
	We consider backscatter energy detection with disjoint decision regions and formulates a \gls{dmtmac}.
	% Specifically, energy detection
	to decode the multiplicative backscatter information under primary uncertainty, and it divides the accumulated energy domain into disjoint decision regions to construct \gls{dmc}
	% , and we design disjoint detection regions to transform the \gls{dcmc} channel to \gls{dmtc}

	Hence, the receiver can semi-coherently (under primary uncertainty) decode RIScatter nodes from the accumulated receive energy, re-encode to recover the exact reflection patterns, and combine those with cascaded \gls{csi} to determine the multipath response and model within the primary channel.
	Such a backscatter-primary decoding enables simultaneous backscatter modulation and dynamic passive beamforming by only one energy comparison and re-encoding during each backscatter (instead of primary) block.

	exploits the properties of active-passive coexisting networks and
	% \emph{Second,} we address the first challenge and propose a practical receiver that exploits the three properties

	It semi-coherently decodes the backscatter information, recovers the exact reflection pattern, and coherently decodes the primary link.
	Thanks to the double modulation effect, once the multiplicative backscatter information is successfully decoded, the contribution is modelled within channel as dynamic passive beamforming rather than re-precoded and cancelled.

	% The semi-coherent backscatter decoding is based on
	% For the semi-coherent backscatter decoding, we
	% Specifically, the backscatter decoding relys on energy
	% we transform the discrete-input continuous-output backscatter channel
	We use energy detection to decode the multiplicative backscatter information under primary uncertainty, and transform the

	we transform the discrete-input continuous-output backscatter channel
	the energy space is divided into disjoint decision regions associated with different hypothese
	% Double modulation enables symbol-level passive precoding such that the
	The multiplicative backscatter information, once successfully decoded, can be modelled as dynamic passive beamforming (rather than being re-precoded and cancelled) and thus the energy detection becomes part of channel training.
	The proposed receiver significantly improves backscatter throughput since it is suitable for arbitrary input distribution and spreading factor.
	This is the first paper to consider practical backscatter-primary decoding scheme at the co-located receiver.

	\emph{Third,} we address the second challenge and

	by considering a specific scenario where multiple co-located/dispersed RIScatter nodes ride over an active point-to-point \gls{miso} transmission to perform backscatter modulation and dynamic passive beamforming towards a nearby user.
	For a given \gls{qos}, we first express the primary and total backscatter rates as functions of \emph{input distribution} at RIScatter nodes, \emph{active beamforming} at the \gls{ap}, and \emph{backscatter decision regions} at the user.
	Then, we characterize the achievable rate region by optimizing those variables under different \gls{qos}.
	Since the original problem is highly non-convex, we propose a suboptimal \gls{bcd} algorithm where the \gls{kkt} input distribution is numerically evaluated by limit of sequences, the active beamforming is iteratively updated by \gls{pgd} accelerated by \gls{bls}, and the decision regions are refined by existing sequential quantization methods for \gls{dmtc}.
	Uniquely, we account for \gls{csi}, \gls{qos}, and backscatter constellation in those optimizations, and the proposed backscatter input distribution design is applicable to more general scenarios.
	This is also the first paper to reveal the importance of backscatter input distribution and decision region designs in active-passive coexisting networks.
